Do I Need to Model You in order to Learn With You?

### 2015-06-05 mutual-modeling.tex
minor
Updated results for study 3, and changed accordingly the discussion

\newcommand{\M}[3]{{\mathcal{M}(#1, #2, #3)}}
\newcommand{\model}[3]{{$\mathcal{M}(#1, #2, #3)$}}
% generic mutual model (gmodel) 
\newcommand{\gmodel}[2]{{$\mathcal{M}(#1, #2)$}}
\newcommand{\refmodel}[2]{{$\mathcal{M}(#1, #2)$}}
\newcommand{\Model}[3]{{$\mathcal{M}^{\circ}(#1, #2, #3)$}}
\newcommand{\gModel}[2]{{$\mathcal{M}^{\circ}(#1, #2)$}}
\newcommand{\Mdeg}[3]{{\mathcal{M}^{\circ}(#1, #2, #3)}}
\newcommand{\gMdeg}[2]{{\mathcal{M}^{\circ}(#1, #2)}}

@@

@@

@@

@@

This leads to two \emph{triangle questions} (Fig.~\ref{mm_triangles}): Do $A$ and
$B$ have the same accuracy when modelling $C$ ($\Delta_2 =
\Delta(\mathcal{M}^{\circ}(A,C,X),
\mathcal{M}^{\circ}(B,C,X))$)? If it is the same, $\mathcal{H}_{2}$ gains over
$\mathcal{H}_{1}$

@@

The aim of this unpublished study was to evaluate the effect of media richness
on \Model{A}{B}{emotions}. The hypothesis was that video communication would
lead to a better \gModel{A}{B} than audio only since emotions often impact on
facial expressions.

@@

The independent variable was the presence or not of video link. The dependent
variable \gModel{A}{B} was measured subjectively from three questionnaires: in the first
one, $A$ described his/her own emotions \gmodel{A}{A}, while in the two other
questionnaire, $A$ described $B$'s and C's emotions. The questionnaire included 18
items (7-point Likert Scale) describing emotions labeled as adjectives:
\emph{anxious}, \emph{enthusiastic}, \emph{agitated}, \emph{proud},
\emph{excited}, \emph{quiet}, \emph{calm}, \emph{stressed}, \emph{bored},
\emph{upset}, \emph{relaxed}, \emph{irritated}, \emph{determined},
\emph{hostile}, \emph{active}, etc. \gmodel{A}{B} was modelled as a vector
of 18 numerical values corresponding to their answers on each
questionnaire items, and \gModel{A}{B} was computed as the correlation between the
two vectors \gmodel{A}{B} and \gmodel{B}{B}.

@@

\paragraph{Study-specific questions} Our hypothesis about media richness is
rejected: the average \Model{A}{B}{emotions} was 0.62 ($SD = 0.28$) with the
audio+video condition and 0.71 ($SD = 0.19$) in the audio condition ($F=3.4, p:
0.065$).

\paragraph{Symmetry question} We computed the absolute differences between
\gModel{A}{B} and \gModel{B}{A} over all subjects (3 values per triad, 20
triads) and averaged them. The average difference in the degrees of accuracy of
symmetrical models is 0.21 ($SD= 0.21$): pairs of subjects tend to assess the
emotions of each other with similar levels of accuracy.

\paragraph{Triangle questions} $\Delta_2$ is computed in a similar way as the average of
the absolute differences between \gModel{A}{C} and \gModel{B}{C} over the 60
subjects. The average difference is 0.17 ($SD= 0.19$), indicating that, inside a
triad, $A$ and $B$ tend to model $C$ in a similar way. For $\Delta_3$, the
average absolute difference between \gModel{C}{A} and \gModel{C}{B} over the 60
subjects is 0.22 ($SD= 0.12$), which means that a given subject tends to model
the two other member of the triad with a similar degree of accuracy, which
supports $\mathcal{H}_{1}$.

@@

In this study, the accuracy of mutual modelling between two peers tend to be
symmetrical, which supports the idea that the quality of modelling is primarily
building upon the social interaction. This is further supported by a low
averaged $\Delta_2$: if, for any $A$ and $B$, $A$ and $B$ generally model $C$
with a similar level of accuracy, it is likely due to the social interaction
induced by $C$ (hypothesis $\mathcal{H}_{2}$).

On the other hand, the low $\Delta_3$ that we also found ($C$ tends to always
model $A$ and $B$ with similar levels of accuracy) tend to contradict this, and
supports the idea that modelling is primarily an individual skill (hypothesis
$\mathcal{H}_{1}$).

@@

\begin{table*}[h!t]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{p{3cm}|p{4cm}|p{4cm}|p{4cm}|p{4cm}|p{4cm}}
\textit{}                     & \textbf{Study 1}                 & \textbf{Study 2}                & \textbf{Study 3}          & \textbf{Study 4}          & \textbf{Study 5}             \\ \hline
\textit{Task}                 & Game in virtual space            & Game in physical space          & Building an argument map  & Building a concept map    & Building a concept map       \\
\textit{Interactions}         & Audio                            & Written                         & Audio / Video             & Audio                     & Audio                        \\
\textit{Shared editor}        & 3D Space                         & 2D Map                          & Concept Map               & Concept Map               & Concept Map                  \\
\textit{Group Size}           & 2                                & 3                               & 3                         & 3                         & 3                            \\
\textit{Duration (mean)}      & 90min                            & 16min                           & 61min                     & 90min                     & 90min                        \\
\textit{Awareness tool}       & Partner's position / orientation & Partner's current/past position & -                         & Partner's concept map     & Partner's scores at pre-test \\
\textit{Dependent var.}       & Team/Ind. performance            & Team/Ind. performance           & Team performance          & Team/Ind. knowledge gains & Team/Ind. knowledge gains    \\
\textit{Independent var.}     & Awareness tool \textit{vs} none  & Awareness tool \textit{vs} none & Audio \textit{vs} Audio+Video & Script \textit{vs} none & Awareness tool \textit{vs} none       \\
\textit{MM°-DV correlation}   & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
\textit{$\Delta_1$}           & +0.38 (significant)              & +0.41 (significant)             & ???                       & +0.26 (near significant)          & +0.05 (ns)              \\
\textit{$\Delta_2$}           &                                  & +0.43 (significant)             & +0.43                     &                           &                              \\
    \textit{$\Delta_3$}       &                                  & +0.30 (significant)             & +0.66                     &                           &                              \\
\textit{$\Delta_4$}           &                                  &                                 &                           & +0.05                     &                             
\end{tabular}
}
\caption{Comparison of the studies}
\label{synthesis_table}
\end{table*}

Coming back to our initial hypotheses, the $\Delta_2$ results in studies 2 and 3
reinforce both $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The \emph{interaction
hypothesis} $\mathcal{H}_{3}$ was that the activity of modelling one's partner
is not simply the result of an individual cognitive effort but overall the
side-effect of the quality of interactions. The \emph{second-order-modelling
hypothesis} $\mathcal{H}_{2}$ stated that, if $A$ is accurate in modelling $B$,
$A$ should notice when $B$ misunderstands $A$ and should be able to repair. A
good modeler would somehow drive the modelling of his or her partner.  This
\emph{drive} will however occur through verbal cues and repairs and is hence not
easy to differentiate from $\mathcal{H}_{3}$.  The contradictory $\Delta_1$
results in studies 4 (to some extent) and 5 weaken the hypothesis.

The triangular questions lead to similar contradictions. The $\Delta_2$
correlation between the \gmodel{A}{C} and \gmodel{B}{C} (studies 2 and 3?)
support $\mathcal{H}_{2}$ if modelling skills were purely individual, there would
be no reason for modelling skills of $A$ and $B$ to be correlated.  The role of $C$ in
\gmodel{A}{C} and \gmodel{B}{C} probably explains this correlation.  Conversely,
the $\Delta_3$ correlation between the \gmodel{A}{B} and \gmodel{A}{C} (studies
3 and 2?) question $\mathcal{H}_{3}$: if the role of the modellee was so
important, it would lead be surprising to have by chance a correlation between
B's and C' ability to drive $A$'s modelling.

In addition, the lack of correlation between \gmodel{A}{A} and \gmodel{A}{B}
($\Delta_2$ in study 4 and study 5) could also contributes to the idea that $A$'s
modelling skills is not the key factor. However, between modelling someone and
modelling oneself, \ie  metacognition, they are so many differences, that we
would hardly conclude from that,

>>>>

\newcommand{\M}[3]{{\mathcal{M}(#1, #2, #3)}}
\newcommand{\model}[3]{{$\mathcal{M}(#1, #2, #3)$}}
% generic mutual model (gmodel) 
\newcommand{\gmodel}[2]{{$\mathcal{M}(#1, #2)$}}
\newcommand{\Gmodel}[2]{{\mathcal{M}(#1, #2)}}
\newcommand{\refmodel}[2]{{$\mathcal{M}(#1, #2)$}}
\newcommand{\Model}[3]{{$\mathcal{M}^{\circ}(#1, #2, #3)$}}
\newcommand{\gModel}[2]{{$\mathcal{M}^{\circ}(#1, #2)$}}
\newcommand{\Mdeg}[3]{{\mathcal{M}^{\circ}(#1, #2, #3)}}
\newcommand{\gMdeg}[2]{{\mathcal{M}^{\circ}(#1, #2)}}

@@

\usepackage{mathtools}
\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
\newcommand*\mean[1]{\overline{#1}}

@@

\paragraph{The symmetry question}

@@

\paragraph{The triangle questions}

@@

This leads to two \emph{triangle questions} (Fig.~\ref{mm_triangles}): Do $A$
and $B$ have the same accuracy when modelling $C$ ($\Delta_2 =
\Delta(\mathcal{M}^{\circ}(A,C,X), \mathcal{M}^{\circ}(B,C,X))$)? If it is the
same, $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$ gains over $\mathcal{H}_{1}$

@@

The aim of this unpublished\footnote{The data and statistical analyses of this
study are available online:
\url{https://github.com/chili-epfl/mutual-modelling-adoption-homosexuality-study}.}
study was to evaluate the effect of media richness on \Model{A}{B}{emotions}.
The hypothesis was that video communication would lead to a better \gModel{A}{B}
than audio only since emotions often impact on facial expressions.

@@

The independent variable was the presence or not of a video link. The dependent
variable \gModel{A}{B} was measured subjectively from three questionnaires: in
the first one, $A$ described his/her own emotions \gmodel{A}{A}, while in the
two other questionnaire, $A$ described $B$'s and $C$'s emotions. The questionnaire
included 18 items (7-point Likert Scale) describing emotions labeled as
adjectives: \emph{anxious}, \emph{enthusiastic}, \emph{agitated}, \emph{proud},
\emph{excited}, \emph{quiet}, \emph{calm}, \emph{stressed}, \emph{bored},
\emph{upset}, \emph{relaxed}, \emph{irritated}, \emph{determined},
\emph{hostile}, \emph{active}, etc. \gmodel{A}{B} was modelled as a vector of 18
numerical values corresponding to their answers on each questionnaire items, and
\gModel{A}{B} was computed as the distance between the two vectors
\gmodel{A}{B} and \gmodel{B}{B}:

\[
    \Mdeg{A}{B}{emotions} = \frac{\displaystyle\sum_{emotions} \abs{\M{A}{B}{e} -
    \M{B}{B}{e}}}{18}
\]

The smaller the score, the \emph{more} accurate the model.

@@

\paragraph{Study-specific questions} Our hypothesis about media richness is
rejected: the average score for \model{A}{B}{emotions} was 1.25 ($SD = 0.53$) in
the audio+video condition and 1.09 ($SD = 0.41$) in the audio alone condition
($t=1.89, df=111, p = 0.062$): the average degree of accuracy of
\model{A}{B}{emotions} was hence \emph{higher} in the audio alone condition.

\paragraph{Symmetry question} We computed the absolute differences $\Delta_1$
between \gModel{A}{B} and \gModel{B}{A} over all pair of subjects within a triad
(3 values per triad, 20 triads), and compared them with the same differences
computed from random gradings (following the same grade distribution as for the
experimental data).

A t-test on the two sets releaved a significantly lower average difference in
the experimental data (mean difference: $0.40$ vs $0.54$ with random gradings,
$t=-3.3, df=60, p = 0.0016$), supporting the symmetry hypothesis.

\paragraph{Triangle questions} $\Delta_2$ is computed in a similar way as the average of
the absolute differences between \gModel{A}{C} and \gModel{B}{C} over the 60
subjects. The average difference is 0.42 ($SD= 0.34$), and if not significantly
different from the same data obtained from random gradings ($t=-1.61, df=60,
p=0.11$).

For $\Delta_3$, the average absolute difference between \gModel{C}{A} and
\gModel{C}{B} over the 60 subjects is 0.40 ($SD= 0.41$) and is significantly
lower than chance ($t=-2.62, df=60, p=0.01$): a given subject tends to model its
two partners with similar degrees of accuracy, which supports $\mathcal{H}_{1}$.

@@

In this study, the accuracy of mutual modelling between two peers tend to be
symmetrical, which would support the idea that the quality of modelling is primarily
building upon the social interaction. This is however not fully supported by the
analysis of $\Delta_2$: $A$ being good at modelling $C$ does not correlate with
$B$ being good at modelling $C$.

On the other hand, $\Delta_3$ shows that a given $A$ tends to
model $B$ and $C$ at similar levels of accuracy: this evidences an individual
skill at modelling others (hypothesis $\mathcal{H}_{1}$).

@@

\begin{table*}[ht!]
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{p{2.6cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}}
\textit{}                     & \textbf{Study 1}                 & \textbf{Study 2}                & \textbf{Study 3}          & \textbf{Study 4}          & \textbf{Study 5}             \\ \hline
\textbf{Task}                 & Game in virtual space            & Game in physical space          & Building an argument map  & Building a concept map    & Building a concept map       \\
\textbf{Interactions}         & Audio                            & Written                         & Audio / Video             & Audio                     & Audio                        \\
\textbf{Shared editor}        & 3D Space                         & 2D Map                          & Concept Map               & Concept Map               & Concept Map                  \\
\textbf{Group Size}           & dyads                            & \textbf{triads}                 & \textbf{triads}           & dyads                     & dyads                            \\
\textbf{Duration (mean)}      & 90min                            & 16min                           & 61min                     & 90min                     & 90min                        \\
\textbf{Awareness tool}       & Partner's position               & Partner's current/past pos.     & -                         & Partner's concept map     & Partner's scores at pre-test \\
\textbf{Dependent var.}       & Team/Ind. performance            & Team/Ind. performance           & Team performance          & Team/Ind. knowledge       & Team/Ind. knowledge          \\
\textbf{Independent var.}     & Awareness tool \textit{vs} none  & Awareness tool \textit{vs} none & Audio \textit{vs} Audio+Video & Script \textit{vs} none & Awareness tool \textit{vs} none       \\
\textbf{Grounding crit.}      & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
$\mathcal{H}_1$               &                                  & supported                       & supported                 &                           &                              \\
$\mathcal{H}_2$               &                                  & supported                       & supported                 &                           &                              \\
$\mathcal{H}_3$               & supported                        & supported                       &                           & supported                 &                              \\
\end{tabular}
}
\caption{Comparison of the studies}
\label{synthesis_table}
\end{table*}

Coming back to our initial hypotheses, the $\Delta_2$ results in study 2
reinforce both $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The \emph{interaction
hypothesis} $\mathcal{H}_{3}$ was that the activity of modelling one's partner
is not simply the result of an individual cognitive effort but overall the
side-effect of the quality of interactions. The \emph{second-order-modelling
hypothesis} $\mathcal{H}_{2}$ stated that, if $A$ is accurate in modelling $B$,
$A$ should notice when $B$ misunderstands $A$ and should be able to repair. A
good modeler would somehow drive the modelling of his or her partner.  This
\emph{drive} will however occur through verbal cues and repairs and is hence not
easy to differentiate from $\mathcal{H}_{3}$. The contradictory $\Delta_2$
analysis in study 3 and $\Delta_1$ results in studies 4 (to some extent) and 5
weaken the hypothesis.

The triangular questions lead to similar contradictions. The $\Delta_2$
correlation between the \gmodel{A}{C} and \gmodel{B}{C} (study 2)
support $\mathcal{H}_{2}$ if modelling skills were purely individual, there would
be no reason for modelling skills of $A$ and $B$ to be correlated.  The role of $C$ in
\gmodel{A}{C} and \gmodel{B}{C} probably explains this correlation.  Conversely,
the $\Delta_3$ correlation between the \gmodel{A}{B} and \gmodel{A}{C} (study
3) question $\mathcal{H}_{3}$: if the role of the modellee was so
important, it would be surprising to have by chance a correlation between
$B$'s and $C$' ability to drive $A$'s modelling.

In addition, the lack of correlation between \gmodel{A}{A} and \gmodel{A}{B}
($\Delta_4$ in study 4 and study 5) could also contributes to the idea that $A$'s
modelling skills is not the key factor. However, between modelling someone and
modelling oneself, \ie  metacognition, they are so many differences, that we
would hardly conclude from that.
